---
title: "Problem Set 10"
author: "Your name and student ID"
date: "Today's date"
output:
  pdf_document: default
---


**Run this chunk of code to load the autograder package!**
```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(testthat)
```

### Instructions 
* Solutions will be released on Tuesday, April 19th.
* This semester, homework assignments are for practice only and will not be turned in for marks.

Helpful hints:

- Every function you need to use was taught during lecture! So you may need to revisit the lecture code to help you along by opening the relevant files on Datahub. Alternatively, you may wish to view the code in the condensed PDFs posted on the course website. Good luck!

- Knit your file early and often to minimize knitting errors! If you copy and paste code for the slides, you are bound to get an error that is hard to diagnose. Typing out the code is the way to smooth knitting! We recommend knitting your file each time after you write a few sentences/add a new code chunk, so you can detect the source of knitting errors more easily. This will save you and the GSIs from frustration! 

- It is good practice to not allow your code to run off the page. To avoid this, have a look at your knitted PDF and ensure all the code fits in the file. If it doesn't look right, go back to your .Rmd file and add spaces (new lines) using the return or enter key so that the code runs onto the next line.

\newpage

You would like to conduct a survey of high school students to determine the proportion who are current e-cigarettes users. Before you conduct your survey, you need to determine how large of a sample size you will need. Suppose that you would like the width of the 95% confidence interval to be 5 percentage points.

**1. Determine the most conservative sample size you would need and assign this value to the object `p1`. Hint: you will need to use a $p^*$ of 0.5.**

```{r}
p1 <- NULL # YOUR CODE HERE
p1

```

```{r}
. = ottr::check("tests/p1.R")
```

\newpage

**2. Next, let's say you've seen a [recent publication](http://annals.org/aim/fullarticle/2698112/prevalence-distribution-e-cigarette-use-among-u-s-adults-behavioral) from the Annals of Internal Medicine that estimated that 9.2% of individuals aged 18 to 24 years old are current e-cigarette users. What is the sample size estimate assuming that $p^*=0.092$?**

```{r}
p2 <- NULL # YOUR CODE HERE
p2

```

```{r}
. = ottr::check("tests/p2.R")
```

\newpage

**3. The publication referenced in the previous question only looked at adults (aged 18+), but observed that the rate of current use was inversely related to age among the population they surveyed. Because of this finding would you suppose that the sample size you estimated in question 2 is too low or too high?**

```{r}
# Uncomment one of the following options:
# p3 <- 'Too low'
# p3 <- 'Too high' 

```

```{r}
. = ottr::check("tests/p3.R")
```

\newpage

Exclusive breastfeeding during the first six months of life is recommended for optimal infant growth and development. Suppose that you conducted a survey of randomly chosen women from California and found that 775 out of 5615 new mothers exclusively breastfed their infants.

Perform all four of the methods discussed in lecture and during lab to create a 95% confidence interval for the proportion of infants exclusively breastfed. 

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(tibble)
```

**4. Construct a 95% confidence interval using the large sample method.**

```{r}
p4 <- NULL # YOUR CODE HERE
p4

```

```{r}
. = ottr::check("tests/p4.R")
```

**5. Construct a 95% confidence interval using the Clopper Pearson (Exact) method.**

```{r}
p5 <- NULL # YOUR CODE HERE
p5

```

```{r}
. = ottr::check("tests/p5.R")
```

**6. Construct a 95% confidence interval using the Wilson Score method with the continuity correction.** 

```{r}
p6 <- NULL # YOUR CODE HERE
p6

```

```{r}
. = ottr::check("tests/p6.R")
```

**7. Construct a 95% confidence interval using the Plus Four method.**

```{r}
p7 <- NULL # YOUR CODE HERE
p7

```

```{r}
. = ottr::check("tests/p7.R")
```

\newpage

**8. Create a plot to compare the confidence interval bounds. If you are stuck, refer back to the example code presented in lab 10.**

```{r}
p8 <- NULL # YOUR CODE HERE
p8

```

\newpage 

**9. Do the methods produce confidence intervals that are basically the same or very different? Why?**

_Type your answer here, replacing this text._

\newpage 

**10. Suppose that in 2010, the rate of exclusive breastfeeding in California was known to be 18.6%. Based on the 95% CIs calculated in questions 4-7, is there evidence against the null hypothesis that the underlying rate is equal to 18.6% in favor of the alternative that the rate is different from 18.6%?**

_Type your answer here, replacing this text._

\newpage 

To confirm your answer to question 10, perform a two-sided hypothesis test and interpret the p-value. 

**11. State the null and alternative hypotheses.**

_Type your answer here, replacing this text._

\newpage

**12. Calculate the test statistic.**

```{r}
p12 <- NULL # YOUR CODE HERE
p12

```

```{r}
. = ottr::check("tests/p12.R")
```

\newpage

**13. Calculate the p-value.**

```{r}
p13 <- NULL # YOUR CODE HERE
p13

```

```{r}
. = ottr::check("tests/p13.R")
```

\newpage

**14. Interpret this p-value in the context of the question.**

_Type your answer here, replacing this text._

\newpage

The quadrivalent HPV vaccine was introduced to Canada in 2007, and was given to girls in Ontario, Canada who were enrolled in grade 8 (13-14 years old). Before 2007, no girls received the vaccine, while in the 4 years after it was introduced nearly 40% of girls received the vaccine each year. One concern was that the vaccine itself would increase promiscuity if the girls felt a false sense of protection, which could thereby increase the prevalence of other sexually transmitted infections (STIs) among vaccinated girls. [This paper](https://www.ncbi.nlm.nih.gov/pubmed/25487660) examines this question using an advanced method called the "regression discontinuity" design which harnesses the abrupt change in vaccination status across cohorts of girls to estimate the causal effect of vaccination against HPV on the occurrence of other STIs.

For the following questions, you only need to read the abstract of the paper. Note that the term "RD" is the difference in risk of STIs between girls exposed and unexposed to HPV vaccination. We can therefore think of this risk difference as the difference between two proportions.

Interpret this result from the abstract: We identified 15/441 (5.9%) cases of pregnancy and sexually transmitted infection and found no evidence that vaccination increased the risk of this composite outcome: RD per 1000 girls -0.61 (95% confidence interval [CI] -10.71 to 9.49). 

**15. In particular, what does -0.61 estimate?**

_Type your answer here, replacing this text._

\newpage

**16. The 95% confidence interval includes 0. What can you conclude about the p-value for a two-sided test of the difference between vaccinated and unvaccinated girls and their risk of sexually transmitted infections?**

_Type your answer here, replacing this text._

\newpage

An allergy to peanuts is increasingly common in Western countries. A randomized controlled trial enrolled infants with a diagnosed peanut sensitivity. Infants were randomized to either avoid peanuts or to consume them regularly until they reached age 5. At the end of the study, 18 out of the 51 randomized to avoid peanuts were tested to be allergic to peanuts. Only 5 out of the 47 randomized to consuming them regularly were tested to be allergic to peanuts. 

**17. Estimate the difference between the two proportions.**

```{r}
p17 <- NULL # YOUR CODE HERE
p17

```

```{r}
. = ottr::check("tests/p17.R")
```

\newpage

**18. Use the plus four method to find a 99% confidence interval for the difference between the two groups. Store the upper and lower bounds into the object `p18`.** 

```{r}
p18 <- NULL # YOUR CODE HERE
p18

```

```{r}
. = ottr::check("tests/p18.R")
```

\newpage

**19. Why would it have been inappropriate to use the large sample method to create a 99% CI?**

_Type your answer here, replacing this text._

\newpage

Perform a two-sided hypothesis test for the difference between the groups. Start by stating the null and alternative hypotheses, then calculate the test statistic, the p-value, and conclude with your interpretation of the p-value.

**20. State the null and alternative hypotheses.**

_Type your answer here, replacing this text._

\newpage 

**21. Calculate the test statistic.**

```{r}
p21 <- NULL # YOUR CODE HERE
p21

```

```{r}
. = ottr::check("tests/p21.R")
```

\newpage

**22. Calculate the p-value.**

```{r}
p22 <- NULL # YOUR CODE HERE
p22

```

```{r}
. = ottr::check("tests/p22.R")
```

\newpage

**23. Interpret the p-value in the context of this question.**

_Type your answer here, replacing this text._

\newpage

**24. Suppose you were testing the hypotheses $H_0: \mu_d = 0$ and $H_a: \mu_d \neq 0$ in a paired design and obtained a p-value of 0.21. Which one of the following could be a possible 95% confidence interval for $\mu_d$?**

```{r}
# Uncomment one of the following choices:
# p24 <- '-2.30 to -0.70'
# p24 <- '-1.20 to 0.90'
# p24 <- '1.50 to 3.80'
# p24 <- '4.50 to 6.90'

```

```{r}
. = ottr::check("tests/p24.R")
```

\newpage 

**25. Suppose you were testing the hypotheses $H_0: \mu_d = 0$ and $H_a: \mu_d \neq 0$ in a paired design and obtain a p-value of 0.02. Also suppose you computed confidence intervals for $\mu_d$. Based on the p-value which of the following are true?**

```{r}
# Uncomment one of the following choices:
# p25 <- 'Both a 95% CI and a 99% CI will contain 0.'
# p25 <- 'A 95% CI will contain 0, but a 99% CI will not.'
# p25 <- 'A 95% CI will not contain 0, but a 99% CI will.'
# p25 <- 'Neither a 95% CI nor a 99% CI interval will contain 0.'

```

```{r}
. = ottr::check("tests/p25.R")
```


## END
