---
title: "Lecture 02:  Begining to work with data"

---

<!-- libraries -->
```{r,include=FALSE,purl=FALSE}
library(knitr) # for include_graphics() 
```

<!-- define default size for R graphics -->
```{r,include=FALSE,purl=FALSE}
outputFormat<-opts_knit$get('rmarkdown.pandoc.to')
if(outputFormat=="beamer"){
  opts_chunk$set(fig.width=6,fig.height=4)
}
```

<!-- define function for external images -->
```{r,include=FALSE,purl=FALSE}
image<-function(ff,ss,cc=NULL,ll=NULL){
  if(is.null(cc)){
    paste('\\centering','\n',
          '\\includegraphics[scale=',ss,']{',ff,'}',
          sep='')
  } else {
    paste('\\centering','\n',
          '\\copyrightbox[b]{',
          '\\includegraphics[scale=',ss,']{',ff,'}',
          '}{\\raggedleft{\\tiny \\href{',ll,'}{',cc,'}}}',
          sep='')    
  }
}
```

<!-- define function for links -->
```{r,include=FALSE,purl=FALSE}
link<-function(tt,ll){
  paste('[\\textcolor[HTML]{ffa328}{\\ul{',tt,'}}]','(',ll,')',sep='')
}
```

```{r load-libraries, echo = F, message=F, warning=FALSE}
library(readr)
library(ggplot2)
library(dplyr)
library(forcats)
```

## So you have some DATA - now what?

### Learning objectives for today:

1. What is a data frame
2. Get the data into R
3. Figure out what's in the dataset
 - Identifying the unit of analysis 
 - Differentiating between the types of variables

4. Manipulate the data frame using the R package `dplyr`'s main functions:
    - `rename()`
    - `select()`
    - `arrange()`
    - `filter()`
    - `mutate()`
    - `group_by()`
    - `summarize()`

### What is a data frame?

- A data frame is a data set. 
- We read data into R from common sources like Excel spreadsheets (.xls or .xlsx), 
    text files (.txt), comma separate value files (.csv), and other formats.
- The simplest format of data contains one row for each individual in the study. 
- The first column of the data identifies the **individual** (perhaps by a name or an ID **variable**). 
- Subsequent columns are **variables** that have been recorded or measured.

### Lake data from Baldi and Moore (B&M)

- Exercise 1.25 from Edition 4 of B&M
- Data from a study of mercury concentration across 53 lakes 
- I've placed these data in my working directory 
- Let's find it there

### readr is a library to import data into R

- To access readr's functions we load the library like this:

```{r}
library(readr)
```

- Click the green arrow to run the code 
- A green rectangle that temporarily appears next to the code shows you that it 
has run.

### `read_csv()` to load the lake data in R
- `read_csv()` is a function from the `readr` library used to import csv files.
- code template: `your_data <- read_csv("pathway_to_data.csv")`
- The `<-` is called the **assignment operator**. It says to save the imported
data into an object called `your_data`.

```{r read-lake-data, message=FALSE}
lake_data <- read_csv("mercury-lake.csv")
```

### Exercise

1. Execute the above code using either the green arrow 
2. Note that the data appears in the Environment pane in the top right. 
   - Notice the number of **observations** and the number of **variables**. 
3. Click the tiny table icon to the right of the `lake_data` in the Environment 
pane to open the **Viewer** tab and inspect the data.


## Describing your data:  what are you working with?

### Four R functions to get to know a dataset

* `head(your_data)`: Shows the first six rows of the supplied dataset
* `dim(your_data)`: Provides the number of rows by the number of columns
* `names(your_data)`: Lists the variable names of the columns in the dataset
* `str(your_data)`: Summarizes the above information and more


```{r head-lake-data}
# notice that if I put a # in front of a line of code it will not run
#head(lake_data)
#dim(lake_data)
#names(lake_data)
str(lake_data)
```

### Unit of analysis
The unit of analysis is the major entity you are working with:

 - Bacteria
 - Laboratory test results
 - Individual People
 - Groups of people (couples, households)
 - Villages
 - Countries
 
 Which function in R lets us know how many units we have?
 

### Type of Variable

- **Categorical** variable: A variable that has grouping levels. Mathematically you can calculate the proportion (%) of individuals in each level of the category.
    - **Nominal** variables: have no underlying order or rank. E.g., hospital ID, HIV status (yes/no variables), race  
    - **Ordinal** variables: can be ordered or ranked. E.g., socio-economic status, BMI categories

- **Quantitative** variable: A continuous, numeric variable that you can perform mathematical operations on. Mathematically, we can you take the median or average of these variables
    - **Discrete** variables: can be counted. E.g., number of brain lesions, number of previous births
    - **Continuous** variables: can be measured precisely, with a ruler or scale. E.g, annual income, blood alcohol content, gestational age at birth

## Participation Question

## dplyr functions for data manipulation

### Using `dplyr` functions for data manipulation {.bigger} 

- `rename()`
- `select()`
- `arrange()`
- `filter()`
- `mutate()`
- `group_by()`
- `summarize()`

### Load the `dplyr` library to access the functions

```{r load-dplyr}
library(dplyr)
```

- These messages mean that some functions (e.g., filter()) share names with 
functions from other libraries. So, when we use `filter()` we will now use the
`dplyr` version because the `stats` library version has been masked.

- essentially "masking" means that a function is covered over, or masked by a more recently loaded function.

### Function 1: rename()

What do you think rename does?

First print the names of the variables:
```{r print-variable-names}
names(lake_data)
```

Run the rename() function and assign it to `lake_data_tidy`:
```{r}
lake_data_tidy <- rename(lake_data, name_of_lake = lakes)
```

### Function 1: rename()
Then reprint the variable names:
```{r}
names(lake_data_tidy)
```

### Function 1: rename() multiple variables at once

You can rename multiple variables at once:

```{r rename-two-variables}
lake_data_tidy <- rename(lake_data, 
                         name_of_lake = lakes, 
                         ph_level = ph)
```

### Code template for rename() function

`new_dataset <- rename(old_dataset, new_name = old_name)`

Another way to write the above code is to use the **pipe** operator: `%>%`

`new_dataset <- old_dataset %>% rename(new_name = old_name)`

The pipe will become very useful in a few slides...

### Function 2: select()

Based on the output below, what do you think `select()` does? 

```{r select-first-three}
smaller_data <- select(lake_data, lakes, ph, chlorophyll)
names(smaller_data)
```

### Function 2: select()

* We use `select()` to select a subset of **variables**. 
* This is very handy if we inherit a large dataset with several variables that
we do not need.

### Function 2: "negative select()"

We can also use "negative `select()`" to deselect variables. Suppose we wanted
to keep all variables except for `age_data`:

```{r deselect-last-variable}
smaller_data_2 <- select(lake_data, - age_data)
names(smaller_data_2)
```

We place a negative sign in front of `age_data` to remove it from the dataset.

### Rewrite using the pipe operator

```{r}
smaller_data <- lake_data %>% select(lakes, ph, chlorophyll)
smaller_data_2 <- lake_data %>% select(- age_data)
```

* Going forward, we will use the pipe operator to write code using any `dplyr` functions
* This is because we can use the pipe to stack many `dplyr` functions in a row

### Function 3: arrange()

What does arrange do? First type `View(lake_data)` to look at the original data. Then 
run the code and examine its output below. What is different?:

```{r arrange-ph}
View(lake_data)
lake_data %>% arrange(ph)
```

### Function 3: arrange() in descending order

```{r arrange-neg-ph}
lake_data %>% arrange(- ph)
```

### Function 3: arrange() by two variables

```{r arrange-age-and-ph}
lake_data %>% arrange(age_data, -ph)
```

### Function 4: mutate()

- **`mutate()` is one of the most useful functions!** 
- It is used to add new variables to the dataset. Suppose that someone told you 
that the number of fish sampled was actually in hundreds, such that 5 is actually 500. You can use mutate to add a new variable to your dataset that is in the hundreds:

```{r mutate-fish-count}
lake_data_new_fish <- lake_data %>% 
  mutate(actual_fish_sampled = number_fish * 100)

lake_data_new_fish
```

### Use `%>%` to append several lines of code together
- We have saved many of new datasets in our environment! 
- If these datasets were larger, they would take up a lot of space. 
- Rather than saving a new dataset each time, we can make successive changes
to one dataset like this:

```{r rename-select-and-mutate}
tidy_lake_data <- lake_data %>% 
  rename(name_of_lake = lakes) %>%
  mutate(actual_fish_sampled = number_fish * 100) %>%
  select(- age_data, - number_fish)
tidy_lake_data
```

- When you see "%>%", say the words "and then...". For example, "Take `lake_data` and then rename `lakes` to name_of_lake, and then mutate..."

### Use `%>%` to "pipe" several lines of code together

```{r rename-select-and-mutate-eval}
tidy_lake_data <- lake_data %>%
  rename(lake_name = lakes) %>%
  mutate(actual_fish_sampled = number_fish * 100) %>%
  select(- age_data, - number_fish)

#tidy_lake_data
```

### Function 5: `filter()`

Filter is another very useful function! What might `filter()` do? 

### Function 5: `filter()`ing on numeric variables
We use filter to select which rows we want to keep in the dataset. Suppose you were only interested in lakes with `ph` levels of 7 or higher.

```{r filter-on-ph}
lake_data_filtered <- lake_data %>% filter(ph > 7)
lake_data_filtered
```


### Function 5: `filter()`ing on character/string variables

Let's try a few more ways to `filter()` the data set since subsetting data is so 
important:

```{r filter-on-recent}
lake_data %>% filter(age_data == "recent")
```

- `==` is read as "is equal to"


### Function 5: `filter()`ing on character/string variables

```{r filter-not-on-recent}
lake_data %>% filter(age_data != "recent") 
```

- `!=` is read as "is not equal to"

### Function 5: `filter()`ing on character/string variables

```{r filter-on-name}
lake_data %>% filter(lakes %in% c("Alligator", "Blue Cypress")) 
```

- `%in%` is the "in" operator. We are selecting rows where the variable `lakes` 
belongs to the specified list.
- The `c()` combines "Alligator" and "Blue Cypress" into a list

### Function 5: multiple `filter()`s at once
```{r filter-multiple}
lake_data %>% filter(ph > 6, chlorophyll > 30)

#this is the same as:
lake_data %>% filter(ph > 6 & chlorophyll > 30)
```

- A comma or the "and" operator (`&`) are equivalent. Here they say, filter the dataset and keep only rows with `ph > 6` AND `chlorophyll > 30`

### Function 5: `filter()` using "or"
```{r filter-multiple-OR}
lake_data %>% filter(ph > 6 | chlorophyll > 30)
```

- `|` is the OR operator. At least one of `ph > 6` or `chlorophyll > 30` needs 
to be true.

### Functions 6 and 7: group_by() and summarize()

Let's execute the following code and see what it does.

```{r group_by-and-mean}
lake_data %>% 
  group_by(age_data) %>% 
  summarize(mean_ph = mean(ph))
```

What happened?

### Functions 6 and 7: group_by() and summarize()

Another one:

```{r group_by-and-mean-sd}
lake_data %>% 
  group_by(age_data) %>% 
  summarize(mean_ph = mean(ph),
            standard_deviation_ph = sd(ph))
```

### Recap: What functions did we use?

1. `library()` to load `readr` and `dplyr`.
2. `read_csv()` to read csv files from a directory.
3. `head()`, `str()`, `dim()`, and `names()` to look at our imported data.
4. `rename()` to rename variables in a data frame.
5. `select()` to select a subset of variables.
6. `arrange()` to sort a dataset according to one or more variables.
7. `mutate()` to create new variables.
8. `filter()` to select a subset of rows.
9. `group_by() and summarize()` to group the data by a categorial variable and 
calculate a statistic.
10. `mean()` and `sd()` to calculate the mean and standard deviation of variables.

### Recap: What operators did we use?

1. Assignment arrow: `<-`: This is our most important operator!
2. Greater than: `>` There are also: 
    - Less than: `<`
    - Greater than or equal to: `>=`, and,
    - Less than or equal to: `<=` 
4. Is equal to: `==`, and `!=` is not equal to
5. `%in%` to select from a list, where the list is created using `c()`, i.e., 
`lakes %in% c("Alligator", "Annie")`

### Reference material: Additional material

- [Data wrangling cheat sheet](https://www.rstudio.com/wp-content/uploads/2015/02/data-wrangling-cheatsheet.pdf)

### How to export from datahub and save onto your own computer

Some of you may want to edit this file in R markdown by adding notes, etc. In that
case, you can make your edits on datahub and save your updated file on the cloud.
You can additionally save your updated file locally on your computer. Here's how to do that:

1. In the File view window, click the checkbox beside the file you'd like to 
export 
2.  click More > Export. 

This will download the file to your computer's downloads folder.

