---
title: "L07:  Relationships between two categorical variables"

---

<!-- libraries -->
```{r,include=FALSE,purl=FALSE}
library(knitr) # for include_graphics() 
library(dplyr)
```


<!-- define function for links -->
```{r,include=FALSE,purl=FALSE}
link<-function(tt,ll){
  paste('[\\textcolor[HTML]{ffa328}{\\ul{',tt,'}}]','(',ll,')',sep='')
}
```


```{r load-libraries, echo = F, message=F,warning=FALSE}
library(tibble)
library(ggplot2)
library(dplyr)
library(forcats)
```


### Learning objectives for today
Today we will focus on how to visualize and quantify relationships between two categorical variables

- Two way tables
  - marginal vs conditional distributions
- Bar graphs
  - side by side
  - stacked
- Simpson's paradox


### Reminder
Categorical variables are just that, categories.

These can be nominal (no underlying order)

or 

ordinal (ordered)

### Two-way tables

- Two-way, or 2X2 (for a table with two columns and two rows)
    - Used to examine the relationship between 2 categorical variables, originally for those with two levels each
- Foundational to epidemiology, because of the types of variables
we are often interested in

Classic 2X2 looks like this:

|Exposure group | Disease | No disease | Row total |
|-------------|--------------|----------------|-----------|
|Exposed       |A            | B              | A+B       |
|Not Exposed   |C           |D           |C+D       |
|Column total | A+C          | B+D           | Total # observations      |


### Example: Lung cancer and smoking

|Group         | Lung Cancer | No Lung Cancer | Row total |
|-------------|--------------|----------------|-----------|
|Smoker       | 12           | 238            | 250       |
|Non-smoker   | 7            | 743            | 750       |
|Column total | 19           | 981            | 1000      |

    
### Marginal distributions

- The marginal distribution of a variable is the one that is **in the margin** 
of the table (i.e., the Row total or the Column total are the two margins of a 
two-way table).  
- The marginal distribution is the distribution for a single categorical variable
- We learned in Ch. 1 how to plot marginal distributions of categorical variables using `geom_bar()`

### Marginal distributions

|Group         | Lung Cancer | No Lung Cancer | Row total |
|-------------|--------------|----------------|-----------|
|Smoker       | 12           | 238            | 250       |
|Non-smoker   | 7            | 743            | 750       |
|Column total | 19           | 981            | 1000      |


- Overall, what % of the population has lung cancer?
    
- Overall, what % of the population are smokers?
   

### Marginal distributions

|Group         | Lung Cancer | No Lung Cancer | Row total |
|-------------|--------------|----------------|-----------|
|Smoker       | 12           | 238            | 250       |
|Non-smoker   | 7            | 743            | 750       |
|Column total | 19           | 981            | 1000      |


- Overall, what % of the population has lung cancer?
    - Answer: 19/1000 = 1.9%
- Overall, what % of the population are smokers?
    - Answer: 250/1000 25% smoking 
- The **marginal** distribution of lung cancer is 1.9% lung cancer, 98.1% no lung cancer.
    
### Conditional distributions

|Group         | Lung Cancer | No Lung Cancer | Row total |
|-------------|--------------|----------------|-----------|
|Smoker       | 12           | 238            | 250       |
|Non-smoker   | 7            | 743            | 750       |
|Column total | 19           | 981            | 1000      |

- The conditional distribution is the distribution of one variable **within** or **conditional on** the 
level of a second variable
- What is the distribution of lung cancer 
**conditional** on the individuals being smokers?

- What is the conditional distribution of lung cancer **given** individuals are non-smoking?
     
### Conditional distributions

|Group         | Lung Cancer | No Lung Cancer | Row total |
|-------------|--------------|----------------|-----------|
|Smoker       | 12           | 238            | 250       |
|Non-smoker   | 7            | 743            | 750       |
|Column total | 19           | 981            | 1000      |


- The conditional distribution of lung cancer **given** smoking is: 
      -  12/250 = 4.8% smokers and 238/250 = 95.2% 
      
- The conditional distribution of lung cancer **given** non-smoking is:
      - 7/750 = 0.9% smokers and 743/750 = 99.1% non-smokers
     
## Visualizing categorical variables in R  
   
### Marginal and Conditional distributions in R    
- We learned in Ch.1 how to plot marginal distributions of categorical variables using `geom_bar()`
- Can we generalize our use of `geom_bar()` to allow us to plot multiple conditional
distributions? I.e., can we show the conditional distribution of lung cancer for 
smokers and non-smokers on the same plot?

### Visualization in R

First, we encode the data to read into R:

```{r encode-table-into-dataframe, message=F}
library(dplyr)
# students, you don't need to know how to do this
two_way_data <- tribble(~ smoking, ~ lung_cancer, ~ percent, ~number,
                     "smoker", "lung cancer", 4.8, 12,
                     "smoker", "no lung cancer", 95.2,238, 
                     "non-smoker", "lung cancer", 0.9, 7,
                     "non-smoker", "no lung cancer", 99.1, 743)
```


### Bar chart for the visualization of marginal distributions

```{r visualize-bar, out.width="80%", echo=FALSE}
ggplot(two_way_data, aes(x = smoking, y=number)) + 
  geom_bar( stat = "identity")  +
  labs(title = "Marginal Distribution of Smoking") + theme_minimal(base_size = 15)
```


### Conditional distributions

"If there is an explanatory-response relationship, compare the conditional 
distribution of the response variable for the separate values of the explanatory
variable."

- This allows you to visualize the distribution of the response variable for
varying levels of the exposure variable. 

### Dodged bar chart for the visualization of conditional distributions
Syntax:

ggplot(two_way_data, aes(x = smoking, y = percent)) + 

  geom_bar(aes(**fill = lung_cance**r), stat = "identity", **position = "dodge"**)  +

  labs(title = "Dodged bar chart") + theme_minimal(base_size = 15)


### Dodged bar chart for the visualization of conditional distributions

```{r visualize-conditional-bar-chart-dodged, out.width="80%", echo=F}
ggplot(two_way_data, aes(x = smoking, y = percent)) + 
  geom_bar(aes(fill = lung_cancer), stat = "identity", position = "dodge")  +
  labs(title = "Dodged bar chart") + theme_minimal(base_size = 15)
```

### Stacked bar chart for the visualization of conditional distributions
ggplot(two_way_data, aes(x = smoking, y = percent)) + 

  geom_bar(aes(**fill = lung_cancer**), stat = "identity", **position = "stack"**) +

  labs(title = "Stacked bar chart") + theme_minimal(base_size = 15)

### Stacked bar chart for the visualization of conditional distributions
```{r visualize-conditional-bar-chart-stacked, out.width = "80%", echo=FALSE}
ggplot(two_way_data, aes(x = smoking, y = percent)) + 
  geom_bar(aes(fill = lung_cancer), stat = "identity", position = "stack") +
  labs(title = "Stacked bar chart") + theme_minimal(base_size = 15)
```

### Visualization of conditional distributions: three levels of response variable

- Plots like the one above make less sense when there are only two levels of 
both of the variables. This is because once you know the percent of lung cancer
among smokers, you also know the percent of non-lung cancer among smokers.
- Here is another example with 3 levels: Shoe support by gender (from ch. 5):

|Group           | Men  | Women | Row total |
|----------------|--------------|-----------|
|Good support    | 94   | 137   | 231       |
|Average support | 1348 | 581   | 1929      |
|Poor support    | 30   | 1182  | 1212      |
|Column total    | 1472 | 1900  | 3372      |

- The question: How does the distribution of support of shoes worn vary between
shoes made for men and women?


### Visualization of conditional distributions: three levels of response variable

Example using shoe support data from Baldi & Moore page 124 of Ed.4

```{r encode-shoe-data, echo=FALSE}
# students, you don't need to know how to do this
shoe_data <- tribble(~ shoe_support, ~ gender, ~ percent,
                       "good",    "men",   94/1472, 
                       "average", "men",   1348/1472,
                       "poor",    "men",   30/1472,
                       "good",    "women", 137/1900,
                       "average", "women", 581/1900,
                       "poor",    "women", 1182/1900)
shoe_data
```

### Visualization of conditional distributions: three levels of response variable

ggplot(shoe_data, aes(x = gender, y = percent)) + 

  geom_bar(stat = "identity", aes(**fill = shoe_support**), **position = "dodge"**) +

  theme_minimal(base_size = 15)


### Visualization of conditional distributions: three levels of response variable

```{r shoe-dodged, out.width = "80%", echo=F}
ggplot(shoe_data, aes(x = gender, y = percent)) + 
  geom_bar(stat = "identity", aes(fill = shoe_support), position = "dodge") +
  theme_minimal(base_size = 15)
```

### Visualization of conditional distributions: three levels of response variable

ggplot(shoe_data, aes(x = gender, y = percent)) + 

  geom_bar(stat = "identity", aes(**fill = shoe_support**), **position = "stack"**) +

  theme_minimal(base_size = 15)


### Visualization of conditional distributions: three levels of response variable

```{r shoe-stacked, out.width = "80%", echo=F}
ggplot(shoe_data, aes(x = gender, y = percent)) + 
  geom_bar(stat = "identity", aes(fill = shoe_support), position = "stack") +
  theme_minimal(base_size = 15)
```

What do we notice here about the ordering of the categories?

### Visualization of conditional distributions: three levels of response variable

Recall from last class we learned how to reorder factor variables that affect
the look of the plot:


shoe_data <- shoe_data %>% 
  mutate(shoe_support = **fct_relevel**(shoe_support, "good", "average", "poor"))

ggplot(shoe_data, aes(x = gender, y = percent)) + 
  geom_bar(stat = "identity", aes(fill = shoe_support), position = "dodge") +
  theme_minimal(base_size = 15)


### Visualization of conditional distributions: three levels of response variable
```{r shoe-dodge-reorder, out.width="80%", echo=FALSE}
library(forcats)
shoe_data <- shoe_data %>% 
  mutate(shoe_support = fct_relevel(shoe_support, "good", "average", "poor"))

ggplot(shoe_data, aes(x = gender, y = percent)) + 
  geom_bar(stat = "identity", aes(fill = shoe_support), position = "dodge") +
  theme_minimal(base_size = 15)
```

### Visualization of conditional distributions: three levels of response variable

Why might we prefer dodged plots to stacked plots?


## Simpson's Paradox

### Simpson's Paradox: Example from Baldi and Moore

- Here is the data presented in your book to illustrate Simpson's paradox. 
- It looks at mortality rates by community and age group for two communities
```{r copy-data-from-book}

simp_data <- tribble(~ age_grp, ~ community, ~ deaths, ~ pop,
                     "0-34", "A", 20, 1000, 
                     "35-64", "A", 120, 3000,
                     "65+", "A", 360, 6000,
                     "all", "A", 500, 10000,
                     "0-34", "B", 180, 6000,
                     "35-64", "B", 150, 3000,
                     "65+", "B", 70, 1000,
                     "all", "B", 400, 10000)
simp_data <- simp_data %>% 
  mutate(death_per_1000 = (deaths/pop) * 1000)
simp_data_no_all <- simp_data %>% filter(age_grp != "all")
```

### Simpson's Paradox Example: Only Conditional data
Plot the mortality rates according to age group and community, linking size of dot to population size 

ggplot(simp_data_no_all, aes(x = age_grp, y = death_per_1000)) + 

  geom_point(aes(**col = community, size = pop**)) + 

  labs(title = "Death rate by age group, community, and population size") +

  theme_minimal(base_size = 15)


### Simpson's Paradox Example: Only Conditional data
```{r simpson-para-1, out.width="50%", echo=FALSE, fig.align='center'}
ggplot(simp_data_no_all, aes(x = age_grp, y = death_per_1000)) + 
  geom_point(aes(col = community, size = pop)) + 
  labs(title = "Death rate by age group, community, and population size") +
  theme_minimal(base_size = 15)
```
- What do we notice about mortality by age groups?
- Which community is larger?
- If someone ask you which community has higher mortality, what would you say?

### Simpson's Paradox Example: with marginal data
```{r simpsons-with-marginal, out.width="50%", fig.align='center', echo=FALSE}
ggplot(simp_data, aes(x = age_grp, y = death_per_1000)) + 
  geom_point(aes(col = community, size = pop)) + 
  labs(title = "Death rate by age group, community, and population size") +
  theme_minimal(base_size = 15)
```

- Notice that the mortality rates for the communities overall show community A
having a higher rate than community B. Why is that?

### Simpson's Paradox

"An association or comparison that holds for all of several groups can reverse
direction when the data are combined to form a single group. This reversal is 
called Simpson's Paradox"


### Simpson's Paradox

- Here are the same data shown using a bar chart 

```{r bar-chart-simpsons-paradox, out.width="50%", echo=FALSE}
#students, don't need to know the scale_fill_manual() function
ggplot(simp_data, aes(x = community, y = death_per_1000)) + 
  geom_bar(aes(fill = age_grp), position = "dodge", stat = "identity", col = "white") +
  theme_minimal(base_size = 15) +
  scale_fill_manual(values = c("#74a9cf", "#0570b0", "#034e7b", "grey"))
  # this scale function allows you to manually set your fill colors
  # colours declared with "#123456" are called hex codes and you can find them online!
```

Which visualization gives you more information?

### Simpson's Paradox Berkeley example

A famous example of Simpson's paradox related to admissions to Berkeley by gender:

Watch: https://www.youtube.com/watch?v=E_ME4P9fQbo

### Recap: Code and concepts

1. `geom_bar(aes(col = var), stat = "identity", position = "dodge")`
2. `geom_bar(aes(col = var), stat = "identity", position = "stack")`
3. Marginal vs conditional distributions
4. Simpson's Paradox


